---
title: 'Analysis for "Taking Control of Regulations"'
author:
- name: Andrew Heiss
  affiliation: Brigham Young University
date: "Last run: `r format(Sys.time(), '%B %e, %Y')`"
output: 
  html_document: 
    highlight: pygments
    theme: sandstone
    code_folding: hide
---

```{r load-libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(magrittr)
library(scales)
library(patchwork)
library(sf)
library(ggrepel)
library(ggstance)
library(pander)
library(here)

pound <- dollar_format(prefix = "£")

source(file.path(here(), "lib", "graphic-functions.R"))
```

# Article 19 budget and staff over time

```{r a19-budget-staff, message=FALSE}
a19_budget <- read_csv(file.path(here(), "data", "a19_budget.csv")) 

a19_budget_long <- a19_budget %>%
  gather(key, value, -Year) %>%
  mutate(type = ifelse(key == "Staff", "Staff", "Budget"))

plot_budget <- ggplot(filter(a19_budget_long, type == "Budget"),
                      aes(x = Year, y = value, color = key)) +
  geom_line(size = 1) +
  labs(x = NULL, y = "Reported amount") +
  scale_y_continuous(labels = pound) +
  scale_color_manual(values = c(ngo_red, ngo_blue_dk), name = NULL) +
  expand_limits(y = 0) +
  theme_ngo()

plot_staff <- ggplot(filter(a19_budget_long, type == "Staff"),
                     aes(x = Year, y = value)) +
  geom_line(size = 1, color = ngo_orange) +
  labs(x = NULL, y = "Full time employees") +
  expand_limits(y = 0) +
  theme_ngo()

plot_budget_staff <- plot_budget + plot_staff + 
  plot_layout(ncol = 1, heights = c(1, 0.5))

plot_budget_staff %T>% 
  print() %T>%
  ggsave(., filename = file.path(here(), "output", "a19-budget-staff.pdf"),
         width = 4.5, height = 3, units = "in", device = cairo_pdf) %>% 
  ggsave(., filename = file.path(here(), "output", "a19-budget-staff.png"),
         width = 4.5, height = 3, units = "in", type = "cairo", dpi = 300)
```

\

Average income per year: `r a19_budget$Income %>% mean() %>% pound()`

Annual income, expenses, and surplus:

```{r a19-budget-staff-tbl, results="asis"}
a19_budget %>%
  mutate(Surplus = Income - Expenses) %>%
  mutate_at(vars(Income, Expenses, Surplus), funs(pound)) %>% 
  pandoc.table(justify = "lllll")
```


# Regional expenses over time

Absolute expenses:

```{r a19-regional-expenses-abs, message=FALSE}
a19_regional <- read_csv(file.path(here(), "data", "a19_regional_expenses.csv")) %>%
  filter(!(Category %in% c("Governance", "Fundraising", "Law and policy"))) %>%
  mutate(Category = ordered(fct_relevel(Category, "Global"))) %>%
  group_by(Year, Category) %>%
  summarise(Amount = sum(Amount))

plot_regions <- ggplot(a19_regional, aes(x = Year, y = Amount)) +
  geom_line(size = 1, color = ngo_red) +
  labs(x = NULL, y = "Reported expenditures") +
  scale_y_continuous(labels = pound) +
  expand_limits(y = 0) +
  theme_ngo(9) +
  facet_wrap(~ Category)

plot_regions %T>% 
  print() %T>%
  ggsave(., filename = file.path(here(), "output", "a19-regions-expenses-abs.pdf"),
         width = 5.5, height = 2.5, units = "in", device = cairo_pdf) %>% 
  ggsave(., filename = file.path(here(), "output", "a19-regions-expenses-abs.png"),
         width = 5.5, height = 2.5, units = "in", type = "cairo", dpi = 300)
```

\

Percent-based expenses:

```{r a19-regional-expenses-pct}
a19_percent <- a19_regional %>%
  group_by(Year) %>%
  mutate(total = sum(Amount),
         percent = Amount / total)

plot_percent <- ggplot(a19_percent, aes(x = Year, y = percent)) +
  geom_line(size = 1, color = ngo_blue_dk) +
  labs(x = NULL, y = "Percent spent on region") +
  scale_y_continuous(labels = percent) +
  expand_limits(y = 0) +
  theme_ngo(9) +
  facet_wrap(~ Category)

plot_percent %T>% 
  print() %T>%
  ggsave(., filename = file.path(here(), "output", "a19-regions-expenses-pct.pdf"),
         width = 5.5, height = 2.5, units = "in", device = cairo_pdf) %>% 
  ggsave(., filename = file.path(here(), "output", "a19-regions-expenses-pct.png"),
         width = 5.5, height = 2.5, units = "in", type = "cairo", dpi = 300)
```


# Regional offices

```{r a19-regional-offices, warning=FALSE, message=FALSE}
# Robinson projection
projection = 54030

world_shapes <- st_read(file.path(here(), "data", "ne_110m_admin_0_countries",
                                  "ne_110m_admin_0_countries.shp"),
                        quiet = TRUE) %>% 
  filter(ISO_A3 != "ATA")

a19_offices <- read_csv(file.path(here(), "data", "a19_offices.csv")) %>% 
  # Convert lat/long to actual geometry object and convert to correct projection
  st_as_sf(coords = c("long", "lat"), crs = 4326) %>% 
  st_transform(projection) %>% 
  # Extract converted lat/long for use with geom_label_repel
  mutate(long = st_coordinates(.)[, 1], lat = st_coordinates(.)[, 2])

plot_a19_offices <- ggplot() +
  geom_sf(data = world_shapes, size = 0.15, color = "black", fill = "grey90") +
  geom_sf(data = a19_offices) +
  geom_label_repel(data = a19_offices, seed = 1234,
                   aes(x = long, y = lat, label = office, fill = hq), 
                   size = 3, color = "white") +
  coord_sf(crs = st_crs(projection), datum = NA) +
  scale_fill_manual(values = c(ngo_blue_dk, ngo_red), guide = FALSE) +
  theme_ngo_map()

plot_a19_offices %T>% 
  print() %T>%
  ggsave(., filename = file.path(here(), "output", "a19-offices.pdf"),
         width = 5.5, height = 2.5, units = "in", device = cairo_pdf) %>% 
  ggsave(., filename = file.path(here(), "output", "a19-offices.png"),
         width = 5.5, height = 2.5, units = "in", type = "cairo", dpi = 300)
```


# Reports in the Middle East

Middle Eastern countries where Article 19 has issued 10+ reports between 2011 and 2017, based on counts from June 19, 2017 at the [Resources database](https://www.article19.org/resources.php?action=resourcesearch&search=test&lang=en).

```{r a19-me-reports, message=FALSE, fig.width=4.5, fig.height=2}
a19_me <- read_csv(file.path(here(), "data", "a19_me_reports.csv"))  %>%
  arrange(reports) %>%
  mutate(country = fct_inorder(country))

plot_me_reports <- ggplot(a19_me, aes(x = reports, y = country, color = hq)) + 
  geom_pointrangeh(aes(xmin = 0, xmax = reports)) +
  labs(x = "Advocacy letters, reports, statements, and updates", y = NULL) +
  scale_colour_manual(values = c(ngo_blue_dk, ngo_red), guide = FALSE) +
  theme_ngo() + theme(panel.grid.major.y = element_blank())

plot_me_reports %T>% 
  print() %T>%
  ggsave(., filename = file.path(here(), "output", "a19-me-reports.pdf"),
         width = 4.5, height = 2, units = "in", device = cairo_pdf) %>% 
  ggsave(., filename = file.path(here(), "output", "a19-me-reports.png"),
         width = 4.5, height = 2, units = "in", type = "cairo", dpi = 300)
```
