---
title: 'Analysis for "Taking Control of Regulations"'
author:
- name: Andrew Heiss
  affiliation: Brigham Young University
date: "Last run: `r format(Sys.time(), '%B %e, %Y')`"
output: 
  html_document: 
    highlight: pygments
    theme: sandstone
    code_folding: hide
---

```{r load-libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(magrittr)
library(scales)
library(patchwork)
library(sf)
library(ggrepel)
library(ggstance)
library(ggdag)
library(ggraph)
library(pander)
library(countrycode)
library(here)

pound <- dollar_format(prefix = "£")

source(here("lib", "graphic-functions.R"))

# Robinson projection
projection = 54030

world_shapes <- st_read(file.path(here(), "data", "ne_110m_admin_0_countries",
                                  "ne_110m_admin_0_countries.shp"),
                        quiet = TRUE) %>% 
  filter(ISO_A3 != "ATA")
```

# Theory

```{r dag, warning=FALSE, message=FALSE, fig.width=5, fig.height=2.75}
theory_dag <- dagify(AE ~ PF,
                     PF ~ IC + RC,
                     FIC ~ AE,
                     labels = c("AE" = "Advocacy\neffects", 
                                "PF" = "Programmatic\nflexibility",
                                "IC" = "Institutional\nconstraints",
                                "RC" = "Resource\nconfiguration",
                                "FIC" = "Future\ninstitutional\nconstraints"),
                     exposure = "PF",
                     outcome = "AE") %>% 
  tidy_dagitty(layout = "sugiyama")

theory_dag$data <- theory_dag$data %>% 
  mutate(linetype = ifelse(name == "AE", "dotted", "solid"))

plot_dag <- ggplot(theory_dag, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(size = 8) +
  geom_dag_edges(start_cap = circle(4, "mm"),
                 end_cap = circle(4, "mm"),
                 aes(edge_linetype = linetype)) +
  geom_dag_label_repel(aes(label = label), direction = "y", seed = 12345, size = 3) +
  scale_edge_linetype_manual(values = c("dotted", "solid"), guide = FALSE) +
  scale_dag() +
  # Recreate scale_dag()'s expand scale settings
  scale_y_reverse(expand_scale(c(0.1, 0.1))) +
  coord_flip() +
  theme_void()

plot_dag %T>% 
  print() %T>%
  ggsave(., filename = here("output", "theory-dag.pdf"),
         width = 5, height = 2.75, units = "in", device = cairo_pdf) %>% 
  ggsave(., filename = here("output", "theory-dag.png"),
         width = 5, height = 2.75, units = "in", type = "cairo", dpi = 300)
```


# CIVICUS restrictions

```{r civicus-data, warning=FALSE, message=FALSE}
# Data from the Tableau visualizations at https://monitor.civicus.org/findings/
civicus <- read_csv(here("data", "civicus_monitor_2017.csv"),
                    na = "Null") %>%
  mutate(Population = as.double(Population),  # integers can't handle world population
         Rating = factor(Rating, levels = c("Open", "Narrowed", "Obstructed", 
                                            "Repressed", "Closed"), 
                         ordered = TRUE),
         iso3 = countrycode(Country, "country.name", "iso3c"))
```

```{r civicus-population}
civicus_population <- civicus %>%
  group_by(Rating) %>%
  summarise(pop_per_rating = sum(Population),
            perc_per_rating = pop_per_rating / sum(civicus$Population)) %>% 
  mutate(open = Rating == "Open")

plot_civicus_population <- ggplot(civicus_population, 
                                  aes(x = perc_per_rating, y = fct_rev(Rating),
                                      color = Rating)) + 
  geom_pointrangeh(aes(xmin = 0, xmax = perc_per_rating)) +
  labs(x = "Percent of world population", y = NULL) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  scale_colour_manual(values = c(ngo_blue_dk, ngo_blue, ngo_yellow, 
                               ngo_orange, ngo_red), 
                      guide = FALSE) +
  theme_ngo() + theme(panel.grid.major.y = element_blank())

plot_civicus_population %T>% 
  print() %T>%
  ggsave(., filename = here("output", "civicus-population.pdf"),
         width = 4.5, height = 2, units = "in", device = cairo_pdf) %>% 
  ggsave(., filename = here("output", "civicus-population.png"),
         width = 4.5, height = 2, units = "in", type = "cairo", dpi = 300)
```

```{r civicus-map}
map_with_civicus <- world_shapes %>% 
  # Fix some Natural Earth ISO weirdness
  mutate(ISO_A3 = ifelse(ISO_A3 == "-99", as.character(ISO_A3_EH), as.character(ISO_A3))) %>% 
  mutate(ISO_A3 = case_when(
    .$ISO_A3 == "GRL" ~ "DNK",
    .$NAME == "Norway" ~ "NOR",
    TRUE ~ ISO_A3
  )) %>% 
  left_join(civicus, by = c("ISO_A3" = "iso3"))

plot_civicus_map <- ggplot() +
  geom_sf(data = map_with_civicus, aes(fill = Rating), size = 0.15, color = "black") +
  coord_sf(crs = st_crs(projection), datum = NA) +
  scale_fill_manual(values = c(ngo_blue_dk, ngo_blue, ngo_yellow, 
                               ngo_orange, ngo_red), 
                    na.translate = FALSE, name = "Civic space") +
  theme_ngo_map() + theme(legend.key.size = unit(0.7, "lines"))

plot_civicus_map %T>% 
  print() %T>%
  ggsave(., filename = here("output", "civicus-map.pdf"),
         width = 5.5, height = 3, units = "in", device = cairo_pdf) %>% 
  ggsave(., filename = here("output", "civicus-map.png"),
         width = 5.5, height = 3, units = "in", type = "cairo", dpi = 300)
```


# Article 19 

## Budget and staff over time

```{r a19-budget-staff, message=FALSE}
a19_budget <- read_csv(here("data", "a19_budget.csv")) 

a19_budget_long <- a19_budget %>%
  gather(key, value, -Year) %>%
  mutate(type = ifelse(key == "Staff", "Staff", "Budget"))

plot_budget <- ggplot(filter(a19_budget_long, type == "Budget"),
                      aes(x = Year, y = value, color = key)) +
  geom_line(size = 1) +
  labs(x = NULL, y = "Reported amount") +
  scale_y_continuous(labels = pound) +
  scale_color_manual(values = c(ngo_red, ngo_blue_dk), name = NULL) +
  expand_limits(y = 0) +
  theme_ngo()

plot_staff <- ggplot(filter(a19_budget_long, type == "Staff"),
                     aes(x = Year, y = value)) +
  geom_line(size = 1, color = ngo_orange) +
  labs(x = NULL, y = "Full time employees") +
  expand_limits(y = 0) +
  theme_ngo()

plot_budget_staff <- plot_budget + plot_staff + 
  plot_layout(ncol = 1, heights = c(1, 0.5))

plot_budget_staff %T>% 
  print() %T>%
  ggsave(., filename = here("output", "a19-budget-staff.pdf"),
         width = 4.5, height = 3, units = "in", device = cairo_pdf) %>% 
  ggsave(., filename = here("output", "a19-budget-staff.png"),
         width = 4.5, height = 3, units = "in", type = "cairo", dpi = 300)
```

\

Average income per year: `r a19_budget$Income %>% mean() %>% pound()`

Annual income, expenses, and surplus:

```{r a19-budget-staff-tbl, results="asis"}
a19_budget %>%
  mutate(Surplus = Income - Expenses) %>%
  mutate_at(vars(Income, Expenses, Surplus), pound) %>% 
  pandoc.table(justify = "lllll")
```


## Regional expenses over time

Absolute expenses:

```{r a19-regional-expenses-abs, message=FALSE}
a19_regional <- read_csv(here("data", "a19_regional_expenses.csv")) %>%
  filter(!(Category %in% c("Governance", "Fundraising", "Law and policy"))) %>%
  mutate(Category = ordered(fct_relevel(Category, "Global"))) %>%
  group_by(Year, Category) %>%
  summarise(Amount = sum(Amount))

plot_regions <- ggplot(a19_regional, aes(x = Year, y = Amount)) +
  geom_line(size = 1, color = ngo_red) +
  labs(x = NULL, y = "Reported expenditures") +
  scale_y_continuous(labels = pound) +
  expand_limits(y = 0) +
  theme_ngo(9) +
  facet_wrap(~ Category)

plot_regions %T>% 
  print() %T>%
  ggsave(., filename = here("output", "a19-regions-expenses-abs.pdf"),
         width = 5.5, height = 2.5, units = "in", device = cairo_pdf) %>% 
  ggsave(., filename = here("output", "a19-regions-expenses-abs.png"),
         width = 5.5, height = 2.5, units = "in", type = "cairo", dpi = 300)
```

\

Percent-based expenses:

```{r a19-regional-expenses-pct}
a19_percent <- a19_regional %>%
  group_by(Year) %>%
  mutate(total = sum(Amount),
         percent = Amount / total)

plot_percent <- ggplot(a19_percent, aes(x = Year, y = percent)) +
  geom_line(size = 1, color = ngo_blue_dk) +
  labs(x = NULL, y = "Percent spent on region") +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  expand_limits(y = 0) +
  theme_ngo(9) +
  facet_wrap(~ Category)

plot_percent %T>% 
  print() %T>%
  ggsave(., filename = here("output", "a19-regions-expenses-pct.pdf"),
         width = 5.5, height = 2.5, units = "in", device = cairo_pdf) %>% 
  ggsave(., filename = here("output", "a19-regions-expenses-pct.png"),
         width = 5.5, height = 2.5, units = "in", type = "cairo", dpi = 300)
```


## Regional offices

```{r a19-regional-offices, warning=FALSE, message=FALSE}
a19_offices <- read_csv(here("data", "a19_offices.csv")) %>% 
  # Convert lat/long to actual geometry object and convert to correct projection
  st_as_sf(coords = c("long", "lat"), crs = 4326) %>% 
  st_transform(projection) %>% 
  # Extract converted lat/long for use with geom_label_repel
  mutate(long = st_coordinates(.)[, 1], lat = st_coordinates(.)[, 2])

plot_a19_offices <- ggplot() +
  geom_sf(data = world_shapes, size = 0.15, color = "black", fill = "grey90") +
  geom_sf(data = a19_offices) +
  geom_label_repel(data = a19_offices, seed = 1234,
                   aes(x = long, y = lat, label = office, fill = hq), 
                   size = 3, color = "white") +
  coord_sf(crs = st_crs(projection), datum = NA) +
  scale_fill_manual(values = c(ngo_blue_dk, ngo_red), guide = FALSE) +
  theme_ngo_map()

plot_a19_offices %T>% 
  print() %T>%
  ggsave(., filename = here("output", "a19-offices.pdf"),
         width = 5.5, height = 2.5, units = "in", device = cairo_pdf) %>% 
  ggsave(., filename = here("output", "a19-offices.png"),
         width = 5.5, height = 2.5, units = "in", type = "cairo", dpi = 300)
```


## Reports in the Middle East

Middle Eastern countries where Article 19 has issued 10+ reports between 2011 and 2017, based on counts from June 19, 2017 at the [Resources database](https://www.article19.org/resources.php?action=resourcesearch&search=test&lang=en).

```{r a19-me-reports, message=FALSE, fig.width=4.5, fig.height=2}
a19_me <- read_csv(here("data", "a19_me_reports.csv"))  %>%
  arrange(reports) %>%
  mutate(country = fct_inorder(country))

plot_me_reports <- ggplot(a19_me, aes(x = reports, y = country, color = hq)) + 
  geom_pointrangeh(aes(xmin = 0, xmax = reports)) +
  labs(x = "Advocacy letters, reports, statements, and updates", y = NULL) +
  scale_colour_manual(values = c(ngo_blue_dk, ngo_red), guide = FALSE) +
  theme_ngo() + theme(panel.grid.major.y = element_blank())

plot_me_reports %T>% 
  print() %T>%
  ggsave(., filename = here("output", "a19-me-reports.pdf"),
         width = 4.5, height = 2, units = "in", device = cairo_pdf) %>% 
  ggsave(., filename = here("output", "a19-me-reports.png"),
         width = 4.5, height = 2, units = "in", type = "cairo", dpi = 300)
```


# AMERA UK/Egypt/International

## Budget and staff over time

```{r amera-budget-staff, message=FALSE}
amera_budget <- read_csv(here("data", "amera_budget.csv")) %>% 
  mutate(cr_total = `Comic Relief` + Uganda,
         cr_percent = cr_total / Income)

amera_budget_long <- amera_budget %>%
  select(Year, Income, Expenses, Staff) %>%
  gather(key, value, -Year) %>%
  mutate(type = ifelse(key == "Staff", "Staff", "Budget"))

plot_budget_amera <- ggplot(filter(amera_budget_long, type == "Budget"),
                            aes(x = Year, y = value, color = key)) +
  geom_line(size = 1) +
  labs(x = NULL, y = "Reported amount") +
  scale_y_continuous(labels = pound) +
  scale_color_manual(values = c(ngo_red, ngo_blue_dk), name = NULL) +
  expand_limits(y = 0) +
  theme_ngo()

plot_budget_cr <- ggplot(amera_budget, aes(x = Year, y = cr_percent)) +
  geom_line(size = 1, colour = ngo_yellow) +
  labs(x = NULL, y = "Percent of income\nfrom Comic Relief") +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  theme_ngo()

plot_staff_amera <- ggplot(filter(amera_budget_long, type == "Staff"),
                           aes(x = Year, y = value)) +
  geom_line(size = 1, color = ngo_orange) +
  labs(x = NULL, y = "Full time employees") +
  expand_limits(y = 0) +
  theme_ngo()

plot_budget_staff_amera <- plot_budget_amera + plot_budget_cr + plot_staff_amera + 
  plot_layout(ncol = 1, heights = c(1, 0.5, 0.5))

plot_budget_staff_amera %T>% 
  print() %T>%
  ggsave(., filename = here("output", "amera-budget-staff.pdf"),
         width = 4.5, height = 3.5, units = "in", device = cairo_pdf) %>% 
  ggsave(., filename = here("output", "amera-budget-staff.png"),
         width = 4.5, height = 3.5, units = "in", type = "cairo", dpi = 300)
```

Average income per year: `r amera_budget$Income %>% mean() %>% pound()`

Annual income, expenses, and surplus:

```{r amera-budget-staff-tbl, results="asis"}
amera_budget %>%
  mutate(Surplus = Income - Expenses) %>%
  mutate_at(vars(Income, Expenses, Surplus), pound) %>% 
  select(Year, Income, Expenses, Staff, Surplus) %>% 
  pandoc.table(justify = "lllll")
```

## Regional partners

```{r amera-regional-partners, warning=FALSE, message=FALSE}
amera_partners <- read_csv(here("data", "amera_partners.csv")) %>% 
  # Convert lat/long to actual geometry object and convert to correct projection
  st_as_sf(coords = c("long", "lat"), crs = 4326) %>% 
  st_transform(projection) %>% 
  # Extract converted lat/long for use with geom_label_repel
  mutate(long = st_coordinates(.)[, 1], lat = st_coordinates(.)[, 2])

plot_amera_offices <- ggplot() +
  geom_sf(data = world_shapes, size = 0.15, color = "black", fill = "grey90") +
  geom_sf(data = amera_partners) +
  geom_label_repel(data = amera_partners, seed = 1234,
                   aes(x = long, y = lat, label = office, fill = hq), 
                   size = 3, color = "white") +
  coord_sf(crs = st_crs(projection), datum = NA) +
  scale_fill_manual(values = c(ngo_blue_dk, ngo_red), guide = FALSE) +
  theme_ngo_map()

plot_amera_offices %T>% 
  print() %T>%
  ggsave(., filename = here("output", "amera-partners.pdf"),
         width = 5.5, height = 2.5, units = "in", device = cairo_pdf) %>% 
  ggsave(., filename = here("output", "amera-partners.png"),
         width = 5.5, height = 2.5, units = "in", type = "cairo", dpi = 300)
```
